Facility and Time Step Discretization Effects in Fuel Cycle Simulation

### 2015-09-09 paper.tex
clean up fleet modeling discussion

\newpage
\begin{abstract}
As nuclear power expands, technical, economic, political, and environmental
analyses of nuclear fuel cycles by simulators increase in importance.  To date,
however, current tools are often fleet-based rather than discrete and privately
\end{abstract}

@@

There are various low level phenomena that have potential to affect the larger
outcome of simulations with respect to facility modeling decisions.

One difference between the fleet and individual modeling is the
\emph{quantized shutdown effect}.  If a fleet-based reactor modeling 3-batch
cores is short a half batch of fuel it will produce $\frac{5}{6}\cdot C$ of
power for that time step where $C$ is the power capacity of a single reactor
of the fleet (e.g. 1000 MWe).  If an individually modeled reactor is short a
half batch of fuel, then it will produce no power that time step.  Time step
duration can also have this effect.  For both fleet and individual reactor
modeling, power goes offline in for full time steps.  For example, if a
3-batch fleet based reactor is short $\frac{1}{2}$ batch of fuel with a 3
month time step, then it will produce $\frac{5}{6} \cdot 3 \text{mo} \cdot C $
of energy.  In contrast, a monthly time step might allow the same reactor to
produce $[\frac{5}{6} \cdot 1 \text{mo} + 2 \text{mo}] \cdot C $ of energy if
there was actually a fuel shortage for only 1 month.

Individually modeled reactors each have their own refueling cycles.  Depending
on how the refueling cycles and outages are staggered, power production can
vary significantly.  For example, if all reactors refuel at the same time,
then all reactors would be offline refueling at the same time.  This is
referred to as the \emph{cycle staggering effect} and does not occur in
fleet-based facility models.  An example of poor cycle staggering is shown in
Figure \ref{fig:sync-cycle}.  This effect is closely related to the
\emph{inventory drawdown effect} described next.

@@

Another difference between fleet-based and individual facility modeling
involves fuel sharing.  Power production by a group of individual reactors for
a fixed amount of fuel shortage may vary depending on how the existing fuel is
distributed.  This is illustrated in Figure \ref{fig:fuel-sharing}.  Being
short 3 batches might mean that one reactor has no batches and isn't operating
or that 3 reactors are missing a single batch.  This will be referred to as
the \emph{fuel sharing effect}.

@@

Another consequence of synchronized reactor cycles is fuel fabrication
capacity idling for an entire cycle (e.g. 18 months), and then receiving
requests for new batches for every reactor all at once.  The reactors draw out
inventory from the fuel fabrication facility together in a large quantity of
orders. In order to avoid running short of inventory for making fresh fuel,
the fabrication facility must keep at enough material on-hand to satisfy all
the reactors at once.  If the reactor cycles were staggered, then a much lower
on-hand inventory would be sufficient to prevent fuel shortages. This is
referred to as the \emph{inventory drawdown effect}.  Increasing time step
duration can also cause similar effects as it reduces the time granularity
over which refueling can occur, resulting in larger impulse drawdown on
inventories. Another component of this effect is that incoming and outgoing
material flows occur "simultaneously" on a given time step. Facilities do not
know about incoming material when they make commitments for outgoing material.
For longer time steps, larger incoming quanta of material are not available
for making offers. In general longer time steps and individual reactor
modeling create a need for larger floating inventory buffers in order to avoid
temporary material shortages.

>>>>

\newpage
\begin{abstract}
TODO: Write this abstract.
\end{abstract}

@@

There are various low level phenomena that have potential to affect the larger
outcome of simulations with respect to facility modeling decisions. Several
effects are described below that result from fundamental differences between
these two modeling paradigms.

One difference between the fleet and individual facility modeling is the
\emph{quantized shutdown effect}.  If a fleet-based reactor modeling 3-batch
cores is short a half batch of fuel it will produce $\frac{5}{6}\cdot C$ of
power for that time step where $C$ is the power capacity of a single reactor
of the fleet (e.g. 1000 MWe).  If an individually modeled reactor is short a
half batch of fuel, then it will produce no power that time step.  Time step
duration can also have this effect.  For both fleet and individual reactor
modeling, power goes offline in full time step increments.  For example, if a
3-batch fleet based reactor is short $\frac{1}{2}$ batch of fuel with a 3
month time step, then it will produce $\frac{5}{6} \cdot 3 \text{mo} \cdot C $
of energy.  In contrast, a monthly time step might allow the same reactor to
produce $[\frac{5}{6} \cdot 1 \text{mo} + 2 \text{mo}] \cdot C $ of energy if
the fuel shortage actually ended after only 1 month.

Individually modeled reactors each have their own refueling cycles.  Depending
on how the refueling cycles and outages are staggered, power production can
vary significantly.  For example, if all reactors refuel at the same time,
then all reactors would be offline at the same time.  This is referred to as
the \emph{cycle staggering effect} and does not occur in fleet-based reactor
models.  An example of poor cycle staggering is shown in Figure
\ref{fig:sync-cycle}.  Poor staggering causes spikes of material supply and
demand that can lead to suboptimal resource utilization.  A fuel fabrication
facility might receive requests for new batches for every reactor all at once.
The reactors draw out inventory from the fuel fabrication facility together in
a large quantity of orders.  If the fabrication facility does not have enough
on hand, some number of reactors will need to wait until enough fuel can be
pumped through fabrication.  Avoiding such constraints would require the
fabrication facility to maintain suboptimally large on-hand inventories as a
contingency.

@@

Another difference between fleet-based and individual facility modeling
involves fuel sharing.  Power production by a group of individual reactors for
a fixed amount of fuel shortage may vary depending on how the existing fuel is
distributed.  This is illustrated in Figure \ref{fig:fuel-sharing}.  A system
being short 3 batches might mean that one reactor has no batches and isn't
operating or that 3 reactors are each missing a single batch.  This is
referred to as the \emph{fuel sharing effect}.

@@

One important phenomenon related to time step duration is the \emph{inventory
drawdown effect}.  Increasing time step duration reduces the frequency over
which refueling can occur, resulting in larger impulse drawdown on
inventories. This by itself is not problematic because this is balanced by
larger inventory top-up quantities.  Another component of this effect is the
simultaneity of incoming and outgoing material flows on a given time step.
Facilities do not know about incoming material when they make commitments for
outgoing material.  For longer time steps, larger incoming quanta of material
are not available for making offers. In general longer time steps and
individual reactor modeling create a need for larger floating inventory
buffers in order to avoid temporary material shortages.
